%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%W  intro.tex          FGA documentation                Christian Sievers
%%
%H  $Id$
%%
%Y  2003
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{Introduction}

\atindex{FGA}{@{\FGA}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Overview}

This manual describes the {\FGA} (*Free Group Algorithms*) package,
a {\GAP} package for computations with finitely generated subgroups of
free groups.

This package allows you to (constructively) test membership, to test
conjugacy, and to compute free generators, the rank, the index,
normalizers, centralizers, and intersections
where the groups involved are finitely generated subgroups of free groups.
%
In addition, it provides a finite presentation for the
automorphism group of a finitely generated free group 
and allows to write any such automorphism as word in
these generators.

See Chapter "Functionality of the FGA Package" for details.

Chapter "Installing and Loading the FGA Package" explains
how to install and load the {\FGA} package.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Implementation and background}

The methods which are used work mainly with inverse finite automata,
a variant of a concept known from theoretical computer science.
An inverse finite automaton is a finite state automaton over a
symmetric alphabet, i.e. one in which every letter has an inverse,
such that each transition between two states for a letter corresponds
to a transition in the opposite direction for the inverse letter.

Most of these techniques are described in chapter 4 of \cite{Sims94},
where the same concept is called coset automaton.
The method to obtain this automaton is called basic coset enumeration,
and in fact it is coset enumeration without defining cosets that
are not important, where a coset <Gg> is called important when there
are words <w> and <v> such that <wv> is reduced and denotes an element
of <G> and <w> denotes an element of <Gg>.

In \cite{BirgetEtAl00}, the connection between finitely generated
subgroups of free groups and inverse finite automata is used to transfer
results about the space complexity of problems concerning inverse finite
automata to analogous results about finitely generated subgroups of free
groups.

Chapter 6 of \cite{Sims94} describes the Reidemeister-Schreier
procedure and a variant called extended coset enumeration which yields
a presentation in the given generators.  The {\FGA} package uses a
variation thereof that might be called extended basic coset
enumeration for its constructive membership test: it leaves out the
part of the algorithm that fills in relations and interprets the
resulting extended coset table differently.

Some word oriented algorithms in the {\FGA} package use basic facts about
free groups that can for example be found in \cite{LyndonSchupp77}.

The presentation of the automorphism groups follows \cite{Neumann33}.
The algorithm for writing an automorphism in the generators works
first at the level of Nielsen generators and uses relations from
\cite{Nielsen}.

The theoretical background for most of this implementation is
explained in \cite{Sievers03}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Integration of the package}

The {\FGA} package mainly installs new methods for operations that are
already known to {\GAP}.  They overlap with methods in the {\GAP}
library in the case of groups of finite index.  In this case, {\GAP}s
methods are usually faster, and the {\GAP} package tries to recognize
such cases.

The methods of the {\FGA} package are only applicable when the groups
involved know they are finitely generated.  This may not always be the
case for groups that were not created by methods of the {\FGA}
package.  In such a case you may first call `GeneratorsOfGroup'.

Some of these problems were already sorted out since release 1.0.
Please inform the package author of any remaining problems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E
